MCTS Applied to Co-operative Problems

### 2015-08-06 Report/response.tex

``\emph{Doors} are rendered in Blue when closed, and are invisible when open.''
\paragraph*{\textit{It is unclear if 'Door' closes again when agent on button moves on (or, alternatively, if walking onto a 'Button' effectively destroy the 'Door').}}
We have improved the definition for \emph{Door} in order to make it clearer that the \emph{Door} closes again when the \emph{Buttons} are no longer activated.

``\emph{Doors -} Doors can either be open or closed. When open, the door is passable. When closed, the door is impassable. Doors are open whilst an \emph{Agent} is activating a linked \emph{Button}. If the \emph{Agent} stops activating the \emph{Button} the \emph{Door} will close.''
\paragraph*{\textit{Since 'Agents' are described as dark yellow, and since there are only two shades of yellow used, it would be wiser to describe a 'Goal' as rendered in bright yellow.}}
We have reworded the definition of \emph{Goal} to better describe the colour in relation to \emph{Agents}.

``\emph{Agents -} Agents are the moving objects that can activate buttons and the goal. \emph{Agents} are rendered in dark Yellow''
\paragraph*{\textit{the sentence "Agent 1 reached an item like this -Agent 2 would seemingly gain some purpose and typically head for the other items." is poorly worded and not very informative. It is better explained in the sentences below. Either remove this sentence or reword it.}}
Removed.

\paragraph*{\textit{it would be a good idea to give the maps meaningful names.}}
This has been done.

``\begin{itemize}
\item{\emph{SingleDoor} This map is the same as the map in \ref{InitialState}}
\item{\emph{Pathfinding} This map is the same as level1.txt but without the \emph{Doors} and \emph{Buttons}. This was to test the \emph{Agent's} ability to solve the simplest problem.}
\item{\emph{SymmetricSingleDoor} This map was a modification of the level1.txt - making the map more symmetric and the action parts closer together.}
\item{\emph{ExtendedSide} This map was an extension of level1.txt with a second \emph{Goal} and some extra walls in the game. Spreading the action parts away from the door and buttons was the main purpose of this one}
\item{\emph{SideBySide} This map was a modification of level1.txt - putting a second \emph{Goal} in and making the map partly mirrored. Each \emph{Agent} began in a separate room to each other instead of together.}
\item{\emph{Airlock} This map was a new map, designed to give wildly different roles to each AI - the top Agent would have to travel through two doors (Like an airlock) that the bottom agent had access to the linked \emph{Buttons}. Then the top \emph{Agent} could collect the reward and allow the bottom \emph{Agent} access to the \emph{Goal}.}
\item{\emph{Butterfly} This map was a new map - with two rooms each with a \emph{Goal} and a door. This was designed to force each \emph{Agent} to be let into the \emph{Goal} room, and out of it again.}
\end{itemize}''

\paragraph*{\textit{The MAGA is not explained very well and seems hastily put together. I would say explain this better but given how poorly it does and how badly optimized it seems to be, maybe the authors should get rid of this player altogether. If this work is resubmitted in future I would recommend creating a better thought out version of this player. For now the better VMAGA should stand alone.
}}
The explanations behind the two agents have been improved, and proper optimisation is a potential candidate for future work in this problem domain. We appreciate the reviewer's comment regarding the poor performance of the algorithm, and it could be possibly eliminated however, we have decided to keep this algorithm in the paper as a comparison, in order to show the improvements obtained when using variable macro-actions (VMAGA).

@@

\paragraph*{\textit{Is the problem domain a real time problem? How long is a tick? 40ms? Do all agents return an action within a tick? Possible 5 actions mean up, down left, right and stay where it is?}}
The definition of the problem domain has been improved with a section describing the particular way that actions are collected and executed as well as a list of each possible action that can be chosen.

``\begin{itemize}
\item{\emph{Left} - This will move the \emph{Agent} one grid square to the left. (-1, 0)}
\item{\emph{Right} - This will move the \emph{Agent} one grid square to the right. (1, 0)}
\item{\emph{Up} - This will move the \emph{Agent} one grid square up. (0, -1)}
\item{\emph{Down} - This will move the \emph{Agent} one grid square down. (0, 1)}
\item{\emph{No-Op} - This will not move the \emph{Agent}. It will remain in the same location as the previous game state.}
\end{itemize}''

\paragraph*{\textit{What is "Uniform Cost applied to Trees"? Is it something similar to the famous "Upper Confidence bound applied to Trees"? If not, please cite something.}}
We have corrected the incorrect definition of UCT from Uniform Cost over Trees to Upper Confidence bound applied to Trees.
\paragraph*{\textit{About MCTS, what are "iteration", "UCT Border", and "rollout border"? My guess is iteration means the number of rollouts,
UCT border means the UCT tree search depth limit, and rollout border means the length of the sequence of rollout. Am I correct? These phrases are different from normal terminology used in MCTS.}}
Reworded to more used terminology.

``The MCTS controller is a simple implementation of UCT - with a fixed number of rollouts, UCT tree search depth limit and rollout border. The rollout border is how far in total, the forward model will be allowed to progress before the game is evaluated.''
\paragraph*{\textit{What is the reward of the rollout? Maybe the raw score?}}
The reward of the rollout was indeed the raw score that the game calculates. No additional heuristics were used as very little information was available to the agents from the forward model.
\paragraph*{\textit{About the results, what is averaged? How many runs are averaged? All agents play in cooperation with other agents including the "keyboard AI"? Is the keyboard AI controlled by human beings? Can we assume it plays perfectly well?}}
The round robin tournament has been better explained, with numbers, at the beginning of the Tournament sub-section.
We have removed the definition for the keyboard AI as it was never used in the experiment and was largely used for manual testing and investigation.

``For the main data collection, we ran a round robin tournament of maps and Agents. Each Agent was paired with all 7 Agents (including a copy of itself) and played 47 games on each of the 7 maps.''
\paragraph*{\textit{Please cite something about (1+1)-ES}}
We have added a citation:

### 2015-08-06 Report/ceec.tex
Expanded the discussion

\usepackage{graphicx}
\usepackage{glossaries}
\usepackage{verbatim}
\usepackage{array}

\graphicspath{{stats-processing/}{stats-processing/data/maps/}}

@@

\subsubsection{Maps}
The 7 maps that were included in the tournament were as follows:
\begin{itemize}
\item{\emph{level1.txt} This map is the same as the map in \ref{InitialState}}
\item{\emph{level1E.txt} This map is the same as level1.txt but without the \emph{Doors} and \emph{Buttons}. This was to test the \emph{Agent's} ability to solve the simplest problem.}
\item{\emph{level2.txt} This map was a modification of the level1.txt - making the map more symmetric and the action parts closer together.}
\item{\emph{level3.txt} This map was an extension of level1.txt with a second \emph{Goal} and some extra walls in the game. Spreading the action parts away from the door and buttons was the main purpose of this one}
\item{\emph{level4.txt} This map was a modification of level1.txt - putting a second \emph{Goal} in and making the map partly mirrored. Each \emph{Agent} began in a separate room to each other instead of together.}
\item{\emph{level5.txt} This map was a new map, designed to give wildly different roles to each AI - the top Agent would have to travel through two doors that the bottom agent had access to the linked \emph{Buttons}. Then the top \emph{Agent} could collect the reward and allow the bottom \emph{Agent} access to the \emph{Goal}.}
\item{\emph{level6.txt} This map was a new map - with two rooms each with a \emph{Goal} and a door. This was designed to force each \emph{Agent} to be let into the \emph{Goal} room, and out of it again.}
\end{itemize}

@@

For the tournament, three parameter sets were chosen for comparison and are shown in Figure \ref{mctsTable}.
\begin{table}[!t]
\begin{center}
\begin{tabular}{|p{2.0cm}|p{2.0cm}|m{1.4cm}|m{1.4cm}|}
\hline
\textbf{Budget} & \textbf{Rollouts} & \textbf{UCT Search Limit} & \textbf{Rollout Border} \\
\hline
Small & 75 & 3 & 15 \\
Medium & 200 & 5 & 30 \\
High & 500 & 10 & 45 \\
\hline
\end{tabular}
\caption{Table of parameters for the 3 \gls{mcts} players}
\label{mctsTable}
\end{center}
\end{table}

\paragraph{Macro Action Genetic Algorithm}
The \gls{ga} algorithm was selected for its simplicity and had Macro Actions added in order to improve the forward search capability as well as the amount of computation time per decision available to it. The \gls{maga} used a population size of 10 and tournament selection of 3. Each candidate was a string of 15 actions - with each action performed 3 times in a row. This meant the \gls{maga} could "see" 45 ticks in the future. The design of the \gls{maga} algorithm was inspired by \textit{Perez et al} \cite{perez2013rolling}
\paragraph{Variable Macro Action Genetic Algorithm}
The Variable Macro Action \gls{ga} was created to try to solve the shortcomings of \gls{maga}. The primary premise was to allow the \gls{ga} to evolve the individual macro action lengths and the length of the overall sequence (number of macro actions). The main technique used was a 1 + 1 Evolutionary Strategy \cite{t.back2000evolutionary-co}. A 1 + 1 ES is a very simple \gls{ga} that maintains a single candidate, and in each iteration a mutation is made and if it is an improvement it is saved as the new best candidate. It is discarded if it is not an improvement. ES's are often much simpler than full gls{ga} algorithms to create, often require much less memory and due to performing less within each iteration is able to use more of the available time budget than a full gls{ga}. The algorithm was bounded by a number of parameters that are shown in Figure \ref{vmagaTable}.

\begin{figure}[h]
\centering
\begin{tabular}{| l | l | l |}
\hline
\textbf{Param} & \textbf{Controls} & \textbf{Value} \\
\hline
minNum & Number of Macro Actions & 3 \\
MaxNum & Number of Macro Actions & 10 \\
minLength & Length of Macro Actions & 1 \\
maxLength & Length of Macro Actions & 5 \\
numChance & Chance of altering Number & 0.25 \\
lengthChance & Chance of altering each length & 0.8 \\
actionChance & Chance of altering each action & 0.75 \\
\hline
\end{tabular}
\caption{Table of parameters for the Variable Macro Action Genetic Algorithm}
\label{vmagaTable}
\end{figure}
These values need extensive tuning to realise the full potential of the algorithm and is a source of possible future work.
\pagebreak
\section{Results}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{level1E-txt-pairs-ticks}
\caption{Average ticks taken to complete level1E.txt when paired only with identical Agents}
\label{1epairedTicks}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{level1-txt-pairs-ticks}
\caption{Average ticks taken to complete level1.txt when paired only with identical Agents}
\label{1pairedTicks}
\end{figure}

@@

\pagebreak
\begin{figure}[!ht]
\centering
\includegraphics[scale=0.5]{level5}
\caption{The initial starting state for level5.txt}
\label{Level5State}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width = \linewidth]{level5-txt-scores}
\caption{Average Score for each AI Agent over Map 5}
\label{avgScoreMap5}
\end{figure}

@@

As seen above, in Figure \ref{avgTicksPairAllMaps}, the medium and high budget implementations provided the two quickest completion times across all maps. The differences shown were significant as well as being the only two agents to complete on average in under 1000 ticks.  Figure \ref{avgScorePairAllMaps} showed that the medium and high budget implementations scored significantly better than all other agents.
\subsubsection{GA}
The two \gls{ga} algorithms did not perform very well in this problem domain.

As seen in Figure \ref{avgScorePairAllMaps} and Figure \ref{avgTicksPairAllMaps}, when tasked with solving the problem domain with another identical agent, neither the \emph{GAController} or \emph{VariGA} performed well at all. The \emph{VariGA} only performed significantly better than either \emph{RandomController} or \emph{GAController} in \emph{level1E.txt}. In all other cases, \emph{VariGA} performed similarly to the standard \emph{GAController}. The ability to mutate the lengths of individual action sequences made the \emph{VariGA} a more flexible pathfinder that the \emph{GAController} but did not aid its ability to solve the co-operative problems present elsewhere in the experiment.
\subsubsection{Random}
The \emph{RandomController} performed poorly in this problem domain. On average, the \emph{RandomController} finished under the 2000 tick limit and scored less than half the maximum on average.
\section{Conclusions}
In this paper, we found that a strong MCTS player can solve simple cooperative problems without requiring communication between agents. We also found that a number of other \gls{ai} techniques experienced difficulty when the problem became cooperative. Further work is required in this area.

>>>>

\usepackage{graphicx}
\usepackage{glossaries}
\usepackage{verbatim}
\usepackage{array}
\usepackage{caption}
\captionsetup{belowskip=12pt,aboveskip=4pt}
\graphicspath{{stats-processing/}{stats-processing/data/maps/}}

@@

\subsubsection{Maps}
The 7 maps that were included in the tournament were as follows:
\begin{itemize}
\item{\emph{SingleDoor} This map is the same as the map in \ref{InitialState}}
\item{\emph{Pathfinding} This map is the same as level1.txt but without the \emph{Doors} and \emph{Buttons}. This was to test the \emph{Agent's} ability to solve the simplest problem.}
\item{\emph{SymmetricSingleDoor} This map was a modification of the level1.txt - making the map more symmetric and the action parts closer together.}
\item{\emph{ExtendedSide} This map was an extension of level1.txt with a second \emph{Goal} and some extra walls in the game. Spreading the action parts away from the door and buttons was the main purpose of this one}
\item{\emph{SideBySide} This map was a modification of level1.txt - putting a second \emph{Goal} in and making the map partly mirrored. Each \emph{Agent} began in a separate room to each other instead of together.}
\item{\emph{Airlock} This map was a new map, designed to give wildly different roles to each AI - the top Agent would have to travel through two doors (Like an airlock) that the bottom agent had access to the linked \emph{Buttons}. Then the top \emph{Agent} could collect the reward and allow the bottom \emph{Agent} access to the \emph{Goal}.}
\item{\emph{Butterfly} This map was a new map - with two rooms each with a \emph{Goal} and a door. This was designed to force each \emph{Agent} to be let into the \emph{Goal} room, and out of it again.}
\end{itemize}

@@

For the tournament, three parameter sets were chosen for comparison and are shown in Table \ref{mctsTable}.
\begin{table}[!t]
\begin{center}
\begin{tabular}{|p{2.0cm}|p{2.0cm}|m{1.4cm}|m{1.4cm}|}
\hline
\textbf{Budget} & \textbf{Rollouts} & \textbf{UCT Search Limit} & \textbf{Rollout Border} \\
\hline
Small & 75 & 3 & 15 \\
Medium & 200 & 5 & 30 \\
High & 500 & 10 & 45 \\
\hline
\end{tabular}
\caption{Table of parameters for the 3 \gls{mcts} players}
\label{mctsTable}
\end{center}
\end{table}

\paragraph{Macro Action Genetic Algorithm}
The \gls{ga} algorithm was selected for its simplicity and had Macro Actions added in order to improve the forward search capability as well as the amount of computation time per decision available to it. The \gls{maga} used a population size of 10 and tournament selection of 3. Each candidate was a string of 15 actions - with each action performed 3 times in a row. This meant the \gls{maga} could "see" 45 ticks in the future. The design of the \gls{maga} algorithm was inspired by \textit{Perez et al} \cite{perez2013rolling}
\paragraph{Variable Macro Action Genetic Algorithm}
The Variable Macro Action \gls{ga} was created to try to solve the shortcomings of \gls{maga}. The primary premise was to allow the \gls{ga} to evolve the individual macro action lengths and the length of the overall sequence (number of macro actions). The main technique used was a 1 + 1 Evolutionary Strategy \cite{t.back2000evolutionary-co}. A 1 + 1 ES is a very simple \gls{ga} that maintains a single candidate, and in each iteration a mutation is made and if it is an improvement it is saved as the new best candidate. It is discarded if it is not an improvement. ES's are often much simpler than full gls{ga} algorithms to create, often require much less memory and due to performing less within each iteration is able to use more of the available time budget than a full \gls{ga}. The algorithm was bounded by a number of parameters that are shown in Table \ref{vmagaTable}.
\begin{table}[!t]
\begin{center}
\begin{tabular}{| l | l | l |}
\hline
\textbf{Param} & \textbf{Controls} & \textbf{Value} \\
\hline
minNum & Number of Macro Actions & 3 \\
MaxNum & Number of Macro Actions & 10 \\
minLength & Length of Macro Actions & 1 \\
maxLength & Length of Macro Actions & 5 \\
numChance & Chance of altering Number & 0.25 \\
lengthChance & Chance of altering each length & 0.8 \\
actionChance & Chance of altering each action & 0.75 \\
\hline
\end{tabular}
\caption{Table of parameters for the Variable Macro Action Genetic Algorithm}
\label{vmagaTable}
\end{center}
\end{table}
These values need extensive tuning to realise the full potential of the algorithm and is a source of possible future work.
\section{Results}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{level1E-txt-pairs-ticks}
\caption{Average ticks taken to complete \emph{Pathfinding} when paired only with identical Agents}
\label{1epairedTicks}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{level1-txt-pairs-ticks}
\caption{Average ticks taken to complete \emph{SingleDoor} when paired only with identical Agents}
\label{1pairedTicks}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.5]{level5}
\caption{The initial starting state for \emph{SideBySide}}
\label{Level5State}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width = \linewidth]{level5-txt-scores}
\caption{Average Score for each AI Agent over \emph{SideBySide}}
\label{avgScoreMap5}
\end{figure}

@@

As seen above, in Figure \ref{avgTicksPairAllMaps}, the medium and high budget implementations provided the two quickest completion times across all maps. The differences shown were significant as well as being the only two agents to complete on average in under 1000 ticks.  Figure \ref{avgScorePairAllMaps} showed that the medium and high budget implementations scored significantly better than all other agents.

One possible reason for \gls{mcts} scoring so well in this problem domain is its use of statistics over hundreds and thousands of simulations to provide it with the ability to act in such a way that it handles all eventualities. Statistically,  in certain situations \gls{mcts} would only see rewards in the tree when it was situated on a \emph{Button}. This tended to cause the \gls{mcts} to travle towards \emph{Buttons}, increasing the likelihood that its own simulations would cause it to be situated on the button. Eventually, most other \gls{ai} \emph{Agents} would cross through the open door.
\subsubsection{GA}
The two \gls{ga} algorithms did not perform very well in this problem domain.

As seen in Figure \ref{avgScorePairAllMaps} and Figure \ref{avgTicksPairAllMaps}, when tasked with solving the problem domain with another identical agent, neither the \emph{GAController} or \emph{VariGA} performed well at all. The \emph{VariGA} only performed significantly better than either \emph{RandomController} or \emph{GAController} in \emph{level1E.txt}. In all other cases, \emph{VariGA} performed similarly to the standard \emph{GAController}. The ability to mutate the lengths of individual action sequences made the \emph{VariGA} a more flexible pathfinder that the \emph{GAController} but did not aid its ability to solve the co-operative problems present elsewhere in the experiment.

The \gls{ga} algorithms don't have the stored tree structure of \gls{mcts} with which to gain the statistical model for what happens when they pursue certain actions. This leads to a seemingly poor performance for the \gls{ga} despite \gls{ga} and \gls{mcts} typically performing equally well in other domains.
\subsubsection{Random}
The \emph{RandomController} performed poorly in this problem domain. On average, the \emph{RandomController} finished under the 2000 tick limit and scored less than half the maximum on average.
\section{Conclusions}
In this paper, we found that a strong MCTS player can solve simple cooperative problems without requiring communication between agents. We also found that a number of other \gls{ai} techniques experienced difficulty when the problem became cooperative. Further work is required in this area.
